Ignition of Fuel Beds by Firebrands

### 2022-05-29 conclusions.tex
Updates to species manuscript
Updated chapter layout

\chapter{Conclusions \& Future Work}
\label{chap:conclusion}
The preceding chapters of this document presented findings from studies evaluating the influence of various parameters on the probability of ignition of fuel beds in wild fires. These studies were conducted to address the specific objectives of this work and to add to the current body of knowledge regarding ignition with the overall goal of facilitating a model for quickly and accurately predicting ignition risk to reduce the number of homes lost to wildfires. The specific objectives of this work are as follows:
        \begin{enumerate}
            \item Determine the effects of particle morphology on ignition propensity
            \item Ascertain ignition dependence on heating location(s), mode, and rate
            \item Identify and quantify the influence of environmental conditions on ignition propensity.
            \item Identify the influence of fuel bed chemical composition on flaming ignition.
        \end{enumerate}
The conclusions of each effort are restated in the following sections followed by a discussion of the implications of the results as they pertain to each specific objective and the overall objective of this work.

\section{Particle Size}
        \begin{enumerate}
            \item 
                Smaller particles ignite more readily in porous beds than larger particles when heat transfer from the heater is primarily through conduction. This was evident by higher ignition probabilities, in general, of the smaller particles for a fixed heater temperature. As particle sizes increase radiant heat transfer becomes more important and fuel beds with larger particles were more likely than smaller particles to ignite at extended times (\textgreater 100\si{\second}) due to the increased importance of radiant ignition. 
            \item
                Douglas-fir plates ignite at times where conduction is the dominant mode of heat transfer (\textless 10\si{\second}) due to the higher thermal conductivity of the solid plates. The ignition probability of plates was the most similar to the larger particle, in particular at lower heater temperatures, due to dispersed heating of the porous fuel bed through radiation and the increased thermal conductivity of the plates creating similar temperature profiles. The rise in ignition probability  over a smaller heater temperature range time with temperature results from more consistent contact between the heater and plate surface.
            \item 
                Heat flux delivered to the fuel bed, when compared to heater temperature, is more indicative of ignition likelihood and ignition time for porous fuel beds. Heat flux is a more significant predictor of ignition because it captures differences in heat transfer modes and particle contact that heater temperature values do not. While this finding is not new, what is novel is that the mixed mode of heating (conduction and radiation) has a significant impact on the flaming ignition of fuel beds.
            \item 
                Consideration of the transport characteristics of pyrolyzate gases near the high temperature source can be important for more fully predicting ignition propensity. A \textit{Da} of ignition, in relation to the measured heat flux and thermal diffusivity of the fuel beds, is a promising relationship for predicting ignition for the porous fuel beds.  
        \end{enumerate}

\section{Wind Speed}
        \begin{enumerate}
            \item
            An increase in wind speed above quiescent conditions reduces the temperature required for the flaming ignition of a fuel bed. For example, an increase in wind speed of 3.5\si{\meter\per\second} from quiescent increases the ignition probability of a fuel bed from under 30\% to roughly 100\%. However, a linear increase in wind speed does not result in a linear increase in ignition probability. Thresholds in wind speed exist above which temperatures required to achieve ignition actually increase. For example, when the wind is oriented 45\si{\degree} from the heater centerline increasing the wind from quiescent to 3.5\si{\meter\per\second} reduces the temperature required for ignition probability by 30\%. However, increasing the wind speed from quiescent to 5.8\si{\meter\per\second} reduces the temperature required for ignition by only 25\%. Presumably these thresholds occur because of reductions in residence time.

@@

\section{Species}
        \begin{enumerate}
            \item Of the six materials tested, only three were able to be ignited within below the 750\si{\celsius} maximum temperature of the firebrand surrogate. Two of the three materials for which ignition was not observed were bark. The higher ignition threshold of the bark materials is attributed to higher lignin concentrations. The resistance to ignition present in these materials suggest that using materials high in lignin in locations at risk for firebrand attack in wildfires may reduce ignitions and subsequent losses of structures.

@@

\section{Ember Interactions}
        \begin{enumerate}
            \item An increase of wind speed from 0.5\si{\meter\per\second} to 5.8\si{\meter\per\second} reduced the temperature required for ignition for all heater configurations. The decrease in ignition temperature required ranged from 20\% to 60\% depending on the heater configuration.

@@

\section{Overall model}
    The overall objective of this work is to identify parameters and processes that control the ignition of a fuel bed when an ember lands on it with an anticipated impact of enabling the creation of a simplified model that may be used to determine the likelihood of ignition quickly. A model was created to predict ignition to evaluate the effectiveness of this work in meeting those goals. The model was then applied to the results of other studies that utilized different fuel bed materials and ignition sources to determine cross-study applicability.

@@

\section{Future Work}
  The results in this work have identified controlling parameters for ignition and provided a framework for creating models with capabilities to predict various ignition scenarios. Significant additional work is needed to refine the proposed framework and explore different ignition parameters. 
        \begin{itemize}
            \item The moisture content of the fuel bed was not considered in this work but is a well-studied parameter that significantly affects ignition. Extension of the experiments is necessary to determine how moisture content fits into the proposed framework. The change in fuel bed bulk density due to moisture may sufficiently characterize the difference, but that remains to be seen.

            \item Additional tests with well-characterized smoldering or flaming embers would significantly enhance the knowledge gained in this work. Typical temperatures of physically combusting firebrands are higher than those capable by the apparatus used in this work~\cite{Fateev2017a}. The higher temperatures may influence the ignition propensity. However, combusting embers are susceptible to heat losses and environmental factors where the firebrand surrogates used in this work are mainly independent of these factors. It is unclear how or if the addition of energy sources that react with the environment will shift the parameters identified to be most influential to ignition. A series of highly instrumented and controlled experiments focused on the coupled interactions between a reacting ember, a reacting fuel bed, and changing environmental conditions. I applaud the patience and perseverance of those who may undertake this endeavor in the future.

            \item Relatively limited data is available regarding the thermal properties of porous biomass media. The lack of available information is likely due to the significant variation in solid biomass material properties and the endless configurations (e.g., porosity and particle orientation) in the natural environment. A series of targeted studies determining the thermal properties of common fuel bed materials would likely increase the understanding of how ignition changes between different materials or even similar materials under various packing conditions.

            \item In addition to the limited data available for thermal properties of fuel bed materials, characterization of the chemical composition of materials and the influence of chemical changes on the pyrolysis and subsequent ignition would provide valuable insight into the differences in ignition between materials and environmental conditions. 
        \end{itemize}

>>>>

\chapter{Conclusions \& Ignition Model}
\label{chap:conclusion}
The preceding chapters of this document presented findings from studies evaluating the influence of various parameters on the probability of ignition of fuel beds in wild fires. The specific objectives of this work are as follows:
        \begin{enumerate}
            \item Determine the effects of particle morphology on ignition propensity
            \item Ascertain ignition dependence on heating location(s), mode, and rate
            \item Identify and quantify the influence of environmental conditions on ignition propensity.
            \item Identify the influence of fuel bed chemical composition on flaming ignition.
        \end{enumerate}
The conclusions of each effort are restated in the following sections followed by a discussion of the implications of the results as they pertain to each specific objective and the overall objective of this work.

\section{Sensitivities of Porous Beds and Plates to Ignition by Firebrands}
        \begin{enumerate}
            \item 
                Smaller particles ignite more readily in porous beds than larger particles when heat transfer from the heater is primarily through conduction. This was evident by higher ignition probabilities, in general, of the smaller particles for a fixed heater temperature. As particle sizes increase radiant heat transfer becomes more important and fuel beds with larger particles were more likely than smaller particles to ignite at extended times (\textgreater 100\si{\second}) due to the increased importance of radiant ignition. 
            \item
                Douglas-fir plates ignite at times where conduction is the dominant mode of heat transfer (\textless 10\si{\second}) due to the higher thermal conductivity of the solid plates. The ignition probability of plates was the most similar to the larger particle, in particular at lower heater temperatures, due to dispersed heating of the porous fuel bed through radiation and the increased thermal conductivity of the plates creating similar temperature profiles. The rise in ignition probability  over a smaller heater temperature range time with temperature results from more consistent contact between the heater and plate surface.
            \item 
                Heat flux delivered to the fuel bed, when compared to heater temperature, is more indicative of ignition likelihood and ignition time for porous fuel beds. Heat flux is a more significant predictor of ignition because it captures differences in heat transfer modes and particle contact that heater temperature values do not. While this finding is not new, what is novel is that the mixed mode of heating (conduction and radiation) has a significant impact on the flaming ignition of fuel beds.
            \item 
                Consideration of the transport characteristics of pyrolyzate gases near the high temperature source can be important for more fully predicting ignition propensity. A \textit{Da} of ignition, in relation to the measured heat flux and thermal diffusivity of the fuel beds, is a promising relationship for predicting ignition for the porous fuel beds.  
        \end{enumerate}

\section{Influence of Wind on Flaming Ignition of Porous Wood Fuel Beds}
        \begin{enumerate}
            \item
            An increase in wind speed above quiescent conditions reduces the temperature required for the flaming ignition of a fuel bed. For example, an increase in wind speed of 3.5\si{\meter\per\second} from quiescent increases the ignition probability of a fuel bed from under 30\% to roughly 100\%. However, a linear increase in wind speed does not result in a linear increase in ignition probability. Thresholds in wind speed exist above which temperatures required to achieve ignition actually increase. For example, when the wind is oriented 45\si{\degree} from the heater centerline increasing the wind from quiescent to 3.5\si{\meter\per\second} reduces the temperature required for ignition probability by 30\%. However, increasing the wind speed from quiescent to 5.8\si{\meter\per\second} reduces the temperature required for ignition by only 25\%. Presumably these thresholds occur because of reductions in residence time.

@@

\section{Effect of Fuel Bed Composition on Flaming Ignition Probability}
        \begin{enumerate}
            \item Of the six materials tested, only three were able to be ignited within below the 750\si{\celsius} maximum temperature of the firebrand surrogate. Two of the three materials for which ignition was not observed were bark. The higher ignition threshold of the bark materials is attributed to higher lignin concentrations. The resistance to ignition present in these materials suggest that using materials high in lignin in locations at risk for firebrand attack in wildfires may reduce ignitions and subsequent losses of structures.

@@

\section{Influence of Multiple Firebrands on the Ignition of Fuel Beds}
        \begin{enumerate}
            \item An increase of wind speed from 0.5\si{\meter\per\second} to 5.8\si{\meter\per\second} reduced the temperature required for ignition for all heater configurations. The decrease in ignition temperature required ranged from 20\% to 60\% depending on the heater configuration.

@@

\section{Ignition Model}
    The overall objective of this work is to identify parameters and processes that control the ignition of a fuel bed when an ember lands on it with an anticipated impact of enabling the creation of a simplified model that may be used to determine the likelihood of ignition quickly. A model was created to predict ignition to evaluate the effectiveness of this work in meeting those goals. The model was then applied to the results of other studies that utilized different fuel bed materials and ignition sources to determine cross-study applicability.

@@
